---
title: Avatar R Tutorial
output: rmarkdown::html_vignette
# output: pdf_document
# always_allow_html: true
vignette: >
  %\VignetteIndexEntry{Avatar Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

This R interface communicates with the avatarization engine. For more information about the concepts and avatarization, checkout our main docs.

## System prerequisites

Make sure your system has `libcurl` properly installed.

On a shell on your Ubuntu/Debian system:

```{bash eval=FALSE, include=TRUE}
apt-get update && apt-get install libcurl4-openssl-dev libssl-dev
```

## Installation

Some packages must be installed before `avatar`.

### R packages dependencies

```{r eval=FALSE, include=TRUE}
# requires libcurl to work (see above)
install.packages(c("tibble", "magrittr", "readr", "httr", "dplyr"))
```

Make sure you're using a recent version of `readr`:

```{r eval=FALSE}
# Make sure you're using a recent version, e.g. 2.1.0
packageVersion("readr")
```

### Install the avatar package

Install the package by pointing to the `.tar.gz` source (replace path below).

```{r eval=FALSE, include=TRUE}
install.packages("~/Downloads/avatar_0.x.x.tar.gz", repos = NULL, type = "source")
```

## Loading the library

Once installed, load the library:

```{r, echo=T, warning=FALSE, message=FALSE}
library(avatar)
```

## Getting help

You can list avatar's vignettes:

```{r eval=FALSE, include=TRUE}
browseVignettes(package = "avatar")
```

You can also get documentation about the main functions:

```{r eval=FALSE, include=TRUE}
help(set_server)
```

## Preliminary steps

First you need to load necessary packages and set the server.

```{r, echo = T, warning=FALSE, message=FALSE}
library(avatar)

# Packages below are not necessary to run avatar but must be installed to run this tutorial
library(knitr)
library(dplyr)
library(ade4)
library(ggplot2)
library(plotly)
library(gridExtra)
library(corrplot)
library(readr)
```

Then configure the API endpoint:

```{r echo = T, results = 'hide'}
# Authorized self signed tls certificate
httr::set_config(httr::config(ssl_verifypeer = 0L))

# Set the server and check connection
# Note: replace localhost with the server's URL
set_server("http://localhost:8000")
```

## Quickstart

This is all you need to run and evaluate and avatarization:

```{r}
# Replace with your server address
set_server("http://localhost:8000")

# Authenticate
authenticate("username", "strong_password")

# Upload the dataset
dataset_id <- upload_dataset(iris)
# Start the avatarization job
job <- start_job(dataset_id, list(k = 20))
# Retrieve the avatars
avatars <- get_avatars(job$id)

# Retrieve the privacy metrics
result <- get_job_result(job$id)
print(result$privacy_metrics$hidden_rate)
```

## Avatarization Step by Step

### Import Dataset

```{r}
# Import data
my_data <- iris

# Display dataframe
kable(head(my_data))

# You can control that R understood the data types for each column during importation:
sapply(my_data, class)
```

If R understood the wrong data type, you need to recast it before avatarization or it will change the results. Here for instance, the variable `Species` could have been coded as `(1, 2, 3)`. In this case R would have casted `Species` as numeric while it is a factor.

### Set parameters

Here's the list of parameters you can use for avatarization. The description for each parameter is available in our main docs.

- `k` (required)
- `column_weights`: default=1 for each variable
- `ncp`: default=5.
- `imputation`: imputation parameters.
  - `k` (default=5): number of neighbors for the knn imputation.
  - `method` (default=`knn`): method used for the imputation. 
  - `training_fraction` (default=1): the fraction of the dataset used to train the knn imputer. 
- `seed`: default=NULL.

```{r}
# Upload your dataset
dataset_id <- upload_dataset(iris)

# Set the parameters you want to use for avatarization
parameters <- list(k = 20, column_weights = list(Sepal.Width = 3), ncp = 3, seed = 42, imputation = list(k = 5))

# You don't need to specify every parameter. Most of them have default values (see above).
parameters <- list(k = 120)
```

### Run avatarization

```{r}
# This launches the avatarization and returns immediately
job <- start_job(dataset_id, parameters)

# You can retrieve the avatarized dataset.
# This call will block until the job is done or a timeout is expired.
# You can call this function as often as you want.
avatars <- get_avatars(job$id)

# Once the avatarization is finished, you will be able to manipulate the avatarized dataset:
# Note that the order of the lines have been shuffled, which means that the link
# between original and avatar individuals cannot be made.
kable(head(avatars))

# You can also retrieve the result metadata, including privacy metrics (see below)
# For the full response, checkout help(get_job_result)
result <- avatar::get_job_result(job$id)
```

You can follow the same process to retrieve the avatars with a higher `k`:

```{r}
parameters_higher_k <- list(k = 120)
job2 <- start_job(dataset_id, parameters_higher_k)
large_k_avatars <- get_avatars(job2$id)
kable(head(large_k_avatars))
```

### Evaluate privacy and utility

You can retrieve the privacy and utility metrics from the result object (see our main docs for details about each metric):

```{r}
result$privacy_metrics$hidden_rate
result$privacy_metrics$local_cloaking
```


### Retrieve unshuffled avatars

⚠ **Caution** - Beware before using the code below ⚠

The unshuffled dataset is not protected as the linkage between the unshuffled avatars dataset and the original data is obvious.

**This dataset contains sensitive data**. You will need to shuffle it in order to make it safe.

```r
# Retrieve your job result
# For the full response, checkout help(get_job_result)
result <- avatar::get_job_result(job$id)
# Download the dataset
columns <- result$sensitive_unshuffled_avatars_datasets$columns
download_url <- result$sensitive_unshuffled_avatars_datasets$download_url
sensitive_unshuffled_avatars <- get_dataset(download_url, columns)
```
